Individual heterogeneity in the functional network topography of medial prefrontal cortex

### 2019-06-18 Manuscript.Rmd
final revisions

  However, strong conclusions about functional colocalization require consideration of individual-level heterogeneity in topographic patterns of brain activity. A recognized limitation of group averaging and meta-analysis is that the functional topography of individual brains can be misaligned and blurred [@Fedorenko2012; @Guntupalli2018a; @Michalka2015; @Tobyne2018; @Wang2015; @Woo2014], exaggerating the apparent overlap across domains. This concern is especially pronounced in mPFC, which is subject to considerable idiosyncratic cortical folding [@Lopez2019; @Mackey2014; @Zilles2013] and inter-subject functional variability [@Mueller2013]. An alternative approach is to focus on analyses at the individual-participant level. Individual-level analyses of fMRI data have identified idiosyncratic, reliable, and valid patterns of functional organization that would be blurred in aggregative estimates [@Gordon2017; @Gratton2018b; @Laumann2015; @Tobyne2018], and subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. Recent work has uncovered fine-grained subdivisions within the DN using both data-driven clustering and individually customized seed-based connectivity analysis [@Braga2017; @Braga2019]. It is therefore possible that the previously reported overlap of DN and valuation effects can be attributed to low effective spatial resolution, and that the organization of mPFC and PCC would be better understood at the individual level. An important first step, and the goal of the present paper, is to quantify the degree of topographic heterogeneity of DN effects within mPFC and PCC in a large sample of individuals. This initial evaluation of variability during rest will provide grounds for individualized topographical estimation of behavioral effects (such as valuation) in the future.

@@

  In this study, we aimed to subdivide mPFC and PCC into individual-specific DN and non-DN communities, and to quantify the degree of topographic heterogeneity in the resulting community structure over time and across individuals. We did this by capitalizing on the respective strengths of meta-analysis and subject-specific analyses of brain networks. We used a data-driven procedure to identify two communities that each spanned both mPFC and PCC in each individual participant. We found that the resulting communities had a consistent topographic layout within PCC (according to a label-agnostic similarity metric), whereas their layout in mPFC was individual-specific and stable across test/re-test. We took advantage of the more consistent configuration within PCC to assign meta-analysis-derived labels to the two communities. Because our data-driven method established correspondence between PCC subregions and mPFC subregions, the labels defined in PCC could then be indexed into the more heterogeneous community structure of mPFC in each individual.

  The outline of our paper is as follows. First, we established a search space by selecting parcels from an established brain atlas [@Glasser2016] that corresponded to previously defined DN and limbic networks on the medial cortical wall [@Yeo2011]. A cortical surface-based meta-analysis of the DN and valuation literatures identified a parcel in PCC that was DN-specific at the aggregate level. We then derived a functional connectivity network of all the surface vertices within the search space for each of `r length(dir(path = './Summary'))` individual resting-state fMRI data sets from the Human Connectome Project [HCP; @VanEssen2012], and used the SP algorithm to subdivide each individual's network into DN and non-DN communities (labeled according to which community included the meta-analytically identified DN-specific parcel in PCC). Focusing on individual vertices in the search space rather than the parcels (as is typical in brain network analyses) allowed us to finely delineate the topographic extent of each community. The resulting communities varied topographically across individuals, while also appearing to follow common organizational principles. Sliding window and test-retest analyses showed that these partitionings were highly similar across scanning days within (but not between) individuals, and that individual-level idiosyncrasy was greater in mPFC. Partitionings obtained from the SP algorithm had higher test-retest reliability than did analogous results from seed-based functional connectivity. Lastly, we describe how the structure of the resulting automatically defined DN and non-DN communities both aligns with and differs from recently proposed scheme for identifying subdivisions within the DN [@Braga2017; @Braga2019]. Our work highlights the usefulness of estimating brain effects at the individual level in mPFC and PCC, and provides a new framework and tool set for future investigations of overlap across cognitive domains.

@@

# Let's sttistically compare stuff
testPCC <- RI_PCC[lower.tri(RI_PCC)]
testPFC <- RI_PFC[lower.tri(RI_PFC)]
Overall_PCCvsPFC_permTest <- permute(testPCC, testPFC, statType = mean, paired = T)
Overall_PCCvsPFC_ES <- cohenD(testPCC, testPFC) 
mPFCvs0_wilcox <- wilcox.test(testPFC) # is the mPFC agreement above chance?
mPFCvs0_ES <- cohenD(testPFC)

@@

```{r Thresholded maps (Fig 5), out.width = "50%", fig.align = "center", fig.cap= "A: Thresholded Fiedler vector maps for subject 100307 (top), and its binarized form (bottom). Subthreshold values effectively formed a third community of high-variability vertices. B: Mean of the binarized maps across all participants, indicating the proportion of DN affiliations per vertex in our sample. Note that 7m-based community label reorientation was used to align the partitionings across individuals in a way that was independent from the community analysis itself. This aggregate map shows the common organizational principle of the DN and non-DN communities, while also showing the high level of variability in mPFC."}

@@

``` {r Partition comparison: halves and runs (Fig. 6), fig.align="center", out.width = "100%", fig.cap = "Left: Similarity matrix for 10 example participants (2 scanning days each), showing pattern agreement across days and subjects for PCC and mPFC separately. The block-diagonal structure is indicative of test-retest reliability across days within an individual. Middle: ratio of within-subject ARI to between-subject mean ARI for all individuals across days suggests idiosyncratic community arrangement for both PCC and mPFC (ratios > 1, solid line), with greater subject-specificity in mPFC. Right: within-to-between subject mean ARI ratios for run-specific partitionings again show greater subject-specific organization for mPFC.", echo=FALSE, message = FALSE}
### code for plots in this fig

>>>>

  However, strong conclusions about functional colocalization require consideration of individual-level heterogeneity in topographic patterns of brain activity. A recognized limitation of group averaging and meta-analysis is that the functional topography of individual brains can be misaligned and blurred [@Fedorenko2012; @Guntupalli2018a; @Michalka2015; @Tobyne2018; @Wang2015; @Woo2014], exaggerating the apparent overlap across domains. This concern is especially pronounced in mPFC, which is subject to considerable idiosyncratic cortical folding [@Lopez2019; @Mackey2014; @Zilles2013] and inter-subject functional variability [@Mueller2013]. An alternative approach is to focus on analyses at the individual-participant level. Individual-level analyses of fMRI data have identified idiosyncratic, reliable, and valid patterns of functional organization that would be blurred in aggregative estimates [@Gordon2017; @Gratton2018b; @Laumann2015; @Tobyne2018], and subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. Recent work has uncovered fine-grained subdivisions within the DN using both data-driven clustering and individually customized seed-based connectivity analysis [@Braga2017; @Braga2019]. It is therefore possible that the previously reported overlap of DN and valuation effects can be attributed to low effective spatial resolution, and that the organization of mPFC and PCC would be better understood at the individual level. An important first step, and the goal of the present paper, is to quantify the degree of topographic heterogeneity of DN effects within mPFC and PCC in a large sample of individuals. This evaluation of variability during rest will provide a frame of reference for individualized topographical estimation of behavioral effects (such as valuation) in the future.

@@

  In this study, we aimed to subdivide mPFC and PCC into individual-specific DN and non-DN communities, and to quantify the degree of topographic heterogeneity in the resulting community structure over time and across individuals. We did this by capitalizing on the respective strengths of meta-analysis and subject-specific analyses of brain networks. We used a data-driven network-analysis procedure to identify two communities that each spanned both mPFC and PCC in each individual participant. We found that the resulting communities had a stereotyped topographic layout within PCC (according to a label-agnostic similarity metric), whereas their layout in mPFC was variable across individuals but stable across test/re-test. We took advantage of the more consistent configuration within PCC to assign meta-analysis-derived labels to the two communities. Because our data-driven method established correspondence between PCC subregions and mPFC subregions, the labels defined in PCC could then be indexed into the more heterogeneous community structure of mPFC in each individual.

  The outline of our paper is as follows. First, we established a search space by selecting parcels from an established brain atlas [@Glasser2016] that corresponded to previously defined DN and limbic networks on the medial cortical wall [@Yeo2011]. A cortical surface-based meta-analysis of the DN and valuation literatures identified a parcel in PCC that was DN-specific at the aggregate level. We then derived a functional connectivity network of all the surface vertices within the search space for each of `r length(dir(path = './Summary'))` individual resting-state fMRI data sets from the Human Connectome Project [HCP; @VanEssen2012], and used the SP algorithm to subdivide each individual's network into DN and non-DN communities (labeled according to which community included the meta-analytically identified DN-specific parcel in PCC). Focusing on individual vertices in the search space rather than the parcels (as is typical in brain network analyses) allowed us to finely delineate the topographic extent of each community. The resulting communities varied topographically across individuals, while also appearing to follow common organizational principles. Sliding window and test-retest analyses showed that these partitionings were highly similar across scanning days within (but not between) individuals, and that individual-level idiosyncrasy was greater in mPFC. Partitionings obtained from the SP algorithm had higher test-retest reliability than did analogous results from seed-based functional connectivity. Lastly, we describe how the structure of the resulting automatically defined DN and non-DN communities both aligns with and differs from a recently proposed scheme for identifying subdivisions within the DN [@Braga2017; @Braga2019]. Our work highlights the usefulness of estimating brain effects at the individual level in mPFC and PCC, and provides a new framework and tool set for future investigations of overlap across cognitive domains.

@@

# Let's sttistically compare stuff
testPCC <- RI_PCC[lower.tri(RI_PCC)]
testPFC <- RI_PFC[lower.tri(RI_PFC)]
Overall_PCCvsPFC_permTest <- permute(testPCC, testPFC, statType = mean, paired = T)
Overall_PCCvsPFC_ES <- cohenD(testPCC, testPFC) 
mPFCvs0_wilcox <- wilcox.test(testPFC) # is the mPFC agreement above chance?
mPFCvs0_ES <- DescTools::CohenD(testPFC)

@@

```{r Thresholded maps (Fig 5), out.width = "50%", fig.align = "center", fig.cap= "A: Thresholded Fiedler vector map for subject 100307 (top), and its binarized form (bottom). Subthreshold values effectively formed a third community of high-variability vertices. B: Mean of the binarized maps across all participants, indicating the proportion of DN affiliations per vertex in our sample. Colors represent PCC-based labels ("DN" versus "non-DN"), which were applied in a subsequent step following the data-driven community-detection analysis and which were necessarily well-aligned in PCC. This aggregate map shows the common organizational principle of the DN and non-DN communities, while also showing the high level of variability in mPFC."}

@@

``` {r Partition comparison: halves and runs (Fig. 6), fig.align="center", out.width = "100%", fig.cap = "Left: Similarity matrix for 10 example participants (2 scanning days each), showing pattern agreement across days and subjects for PCC and mPFC separately. Color scale represents the ARI, which quantifies topographic similarity irrespective of how the communities are labeled. The block-diagonal structure is indicative of test-retest reliability across days within an individual. Middle: ratio of within-subject ARI to between-subject mean ARI for all individuals across days suggests idiosyncratic community arrangement for both PCC and mPFC (ratios > 1, solid line), with greater subject-specificity in mPFC. Right: within-to-between subject mean ARI ratios for run-specific partitionings again show greater subject-specific organization for mPFC.", echo=FALSE, message = FALSE}
### code for plots in this fig
